---
title: "Assessment2SecondAttempt"
author: "BrookeKneebush"
date: '2022-04-13'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(randomForest)
library(skimr)
library(broom)
library(widyr)
library(broom.mixed)

```

```{r}

Thanksgiving_Meals<-read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2018/2018-11-20/thanksgiving_meals.csv'
)

```

## *Brooke Kneebush*
*** 
## **S3032877**
*** 
# Assignment 2
*** 

### Part 1:  formatting RMarkdown document

### 1. Create an Rmarkdown document with webpage as output (same as in setup)

### At the start of the output document include your name in italic font and your student id in bold font as level 2 heading 

### Separate with a solid line 

### Include the title “Assignment 2” as level 1 heading 

### Separate with a solid line 

### List all tasks in the assignment as headings of the third level and include your results (=output) below each task showing your R code. 

### Part 2: Data Wrangling and visualization

### For all tables below, you need to use the RMarkdown functionality to present tables (`kable`).

### 1. Display the first 10 rows of the dataset using `kable()` function (1 marks). 
